On the eletrodynamic of moving bodies

### 2019-05-23 Article_AS.Rmd
Intro complétée

Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of canadian forest are harvested each year and an additionnal loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain ressource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better stuctural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. \

@@

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of lossing ressource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Consequently studies have been set up to understand the genetic sources of phenotypic variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. Wood quality has mainly two components: biologic and economic. The biologic component is the resistance to mortality by vigor and the economic component is the value of the wood for end-use. Several traits were considered for wood quality assessment and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006] because this trait is important for both components of quality. \

@@

Unlike growth that is more controlled by environment than genetic and therefore has low to moderate heritability [@Ivkovich2002a], wood density in white spruce is under strong genetic control with moderate to high heritability [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b, @Beaulieu2006] and this makes this trait attractive for genetic improvement.

However, the timeframe of such improved forest management being long [@Zobel1972] it is essential to take into account the current climate change context [@Gray2016]. With the increasing temperature, the trees are expected to growth faster and thus being of lower quality because of lower density [@corriveau1987; @DeBell1994; @Lindstrom1996; @Dutilleul1998; @Bouriaud2005]. In addition to selecting trees that are prone to produce wood of uniform density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accuralty, with climate change. \

In this study we quantified and compared heteregoneity of wood density from trees of known parents in order to figure out if some individus produce more uniform wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the homogeneity. By comparing two growth sites we tried to characterize the environmental contribution to homogeity. We ultimalty aim to select and reproduce trees families of interest that are both  low-sensitivity to climate change and genetically prone to make homogenous wood.

@@

![Study sites](graph_table/Study.sites.png){width=320px}

>>>>

Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of canadian forest are harvested each year and an additionnal loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain ressource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better stuctural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \

@@

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of lossing ressource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood density is among the key traits as it is related mechanical strength and dimensional stability. Variation in density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. Consequently studies have been set up to understand the genetic sources of phenotypic variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Mullin2011, @Beaulieu2009]. Wood quality has mainly two components: biologic and economic. The biologic component is the resistance to mortality by vigor and the economic component is the value of the wood for end-use. Several traits were considered for wood quality assessment and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006] because this trait is important for both components of quality and strongly correlated with different mechanical and anatomical wood traits [@zobel1989]. \

@@

Unlike growth that is more controlled by environment than genetic and therefore has low to moderate heritability [@Ivkovich2002a], wood density in white spruce is under strong genetic control with moderate to high heritability [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b, @Beaulieu2006] and this makes this trait attractive for genetic improvement. Moreover, Climate is a determining factor for the growth and physiological processes of trees. As a result, climatic conditions strongly affect the properties of wood. In fact, certain climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. In addition, long-term growth analyzes have shown that it may increase due to temperature, atmospheric CO2 content and nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Knowing that density can be negatively correlated with growth in spruce [@zhang1995], it is essential to take this factor into account.

However, the timeframe of such improved forest management being long [@Zobel1972] it is essential to take into account the current climate change context [@Gray2016]. With the increasing temperature, the trees are expected to growth faster and thus being of lower quality because of lower density [@corriveau1987; @DeBell1994; @Lindstrom1996; @Dutilleul1998; @Bouriaud2005]. It is necessary to anticipate the possible environmental variations because more important frequencies water stress or important precipitations are feared with the climatic changes [@IPCC2007; @pachauri2014]. These environmental variations may lead to an increase or decrease in radial growth and thus cause greater variations in the density of the wood. In addition to selecting trees that are prone to produce wood of uniform density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accuralty, with climate change. The heritability of density has been proven in several studies, but there is no evidence to suggest that density variability is hereditary. The heritability of this trait could qualify it for the selection criteria for future breeding programs for producing better quality wood.\

In this study we quantified and compared heteregoneity of wood density from trees of known parents in order to figure out if some individus produce more uniform wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the homogeneity. By comparing two growth sites we tried to characterize the environmental contribution to homogeity. We ultimalty aim to select and reproduce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogenous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

@@

![Study sites](graph_table/Study.sites.png){width=500px; height=500px}
